### 2023-10-23 uci-bike-sharing-dataset/Paper_Summary.rmd
Update Paper_Summary.rmd

---
# This is the YAML header/metadata for the document
title: "Bike Share Chronicles_Unveiling Patterns"
author's: "1) Lakshmi Sreya Rapolu 2) Ashwin Muthuraman 3) Sai Srinivas Lakkoju "
# date: "today"
date: "`r Sys.Date()`" 
# this style requires installing rmdformats package 
output:  
    rmdformats::readthedown:
      toc_float: true
      toc_depth: 3
      number_sections: true
      code_folding: hide
      includes:
        before_body: header.tmphtml
---

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
 knitr::opts_chunk$set(warning = F, results = "markup", message = F)
#knitr::opts_chunk$set(warning = F, results = "hide", message = F)

# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

# Introduction

Bike sharing programs are gaining popularity worldwide because they are environmentally friendly and healthy. Cities are developing bike sharing programs to encourage people to ride bicycles.

Riders can rent bikes from manual or automated stations throughout the city for a set period of time. Riders can typically pick up bikes from one location and return them to another designated location.

Bike sharing programs generate a lot of data, such as travel time, start and end locations, and rider demographics. This data can be combined with other sources of information, such as weather,holiday and season, to learn more about how and when people use bike sharing programs.

## Problem Statement

The goal of this case is to perform Exploratory Data Analysis (EDA) on daily bike rental counts, considering environmental and seasonal variables. This EDA aims to understand the historical usage patterns of a bike-sharing program in Washington, D.C., in relation to weather, environmental factors, and other data. We are interested in exploring the relationships between bike rentals and various factors such as season, temperature, and weather conditions. Our objective is to build insights and understanding from the data

# Tests

```{r}
#Effect of Holiday On Casual Users
casual_holiday = aov(Casual ~ Holiday, data = bike_df)
casual__holidaysummary = summary(casual_holiday)
casual__holidaysummary
```

From the ANOVA above analysis with the "Holiday" variable revealed a highly significant effect. There are statistically significant differences in the response variable between holidays and non-holidays, as indicated by the low p-value (3.2e-05) and the F-value of 17.3. This suggests that the "Holiday" variable has a substantial impact on the response variable.

```{r}
#Effect of Working Day on Registered Users
Registered_workingday = aov(Registered ~ Workingday, data = bike_df)
Registered_workingday = summary(Registered_workingday)
Registered_workingday
```

From the above ANOVA analysis involving the "Workingday" variable indicates a highly significant effect. There are substantial and statistically significant differences in the response variable between working days and non-working days, as evident from the very low p-value (<2e-16) and a substantial F-value of 319. These results suggest that the "Workingday" variable significantly impacts the response variable.

```{r}
#Hour Vs Casual Count

bike_df$Hour <- as.numeric(as.character(bike_df$Hour))
#Because 'Estimated effects may be unbalanced'
# Create a new variable to group hours
bike_df <- bike_df %>%
  mutate(TimeOfDay = case_when(
    Hour >= 6 & Hour < 12 ~ "Morning",
    Hour >= 12 & Hour < 18 ~ "Afternoon",
    TRUE ~ "Evening"
  ))
unique(bike_df$TimeOfDay)
time_of_day_anova <- aov(Casual ~ TimeOfDay, data = bike_df)
time_of_day_summary <- summary(time_of_day_anova)
time_of_day_summary
```

The above ANOVA analysis involving the "TimeOfDay" variable reveals a highly significant effect. There are substantial and statistically significant differences in the response variable across different times of the day. The very low p-value (<2e-16) and a substantial F-value of 2174 indicate that the "TimeOfDay" variable significantly impacts the response variable(Casual User Count).

```{r}
time_of_day_anova_R <- aov(Registered ~ TimeOfDay, data = bike_df)
time_of_day_summary_R <- summary(time_of_day_anova_R)
time_of_day_summary_R
```

The above ANOVA analysis involving the "TimeOfDay" variable reveals a highly significant effect. There are substantial and statistically significant differences in the response variable across different times of the day. The very low p-value (<2e-16) and a substantial F-value of 1047 indicate that the "TimeOfDay" variable significantly impacts the response variable(Registered User Count)

```{r}
#Day Of the week Vs Casual
day_of_the_week_anova_c <- aov(Casual ~ Weekday, data = bike_df)
day_of_the_week_summary_C <- summary(day_of_the_week_anova_c)
day_of_the_week_summary_C
```

The ANOVA analysis involving the "Weekday" variable demonstrates a highly significant effect. There are substantial and statistically significant differences in the response variable across different days of the week. The very low p-value (<2e-16) and a substantial F-value of 294 indicate that the "Weekday" variable significantly impacts the response variable(Casual User count).

```{r}
#Day Of the week Vs Registered

day_of_the_week_anova_r <- aov(Registered ~ Weekday, data = bike_df)
day_of_the_week_summary_r <- summary(day_of_the_week_anova_r)
day_of_the_week_summary_r
```

The ANOVA analysis for the "Weekday" variable indicates a highly significant effect on the response variable. There are substantial and statistically significant differences in the response variable across different days of the week. The very low p-value (<2e-16) and a substantial F-value of 46.1 suggest that the "Weekday" variable significantly impacts the response variable(Registered User Count).

```{r}

```